Long range correlations and their event scale dependence in high multiplicity pp collisions at $\sqrt{s} = 13$~TeV

### 2020-06-12 sec_results.tex
JL and BK lastet

Measurement of the correlations function is shown in Fig.~\ref{fig:PlotCorrMBHMT} for 1.0$<\it{p}_{\rm{T, assoc}},\,\it{p}_{\rm{T, trig}}<$2.0 GeV/\it{c}\rm{} in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV} in the 0-0.1\% (left), and 0-100\% (right) multiplicity class estimated by V0 detector, which covers forward rapidity region. The ridge is clearly seen in high multiplicity class unlike in lower multiplicity classes.

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.4 \textwidth]{./figures/corr1.pdf} }
	\subfigure{ \includegraphics[width=0.4 \textwidth]{./figures/corrmb.pdf} }
	\caption{ Two-dimensional correlations function as function of $\Delta\eta$ and $\Delta\varphi$ in 0-0.1\% (left) and 0-100\% (right) multiplicity class. The interval of transverse momentum of trigger particle and associated particle is 1.0$<\it{p}_{\rm{T}}<$2.0 GeV/\it{c}\rm{}. }
	\label{fig:PlotCorrMBHMT}
\end{figure}

The one-dimensional $\Delta\varphi$ distribution is shown in Fig.~\ref{fig:PlotDeltaPhi} for pairs of particles with various $\it{p}_{\rm{T}}$ intervals in very high multiplicity class. The associated yield per trigger particle is compared with CMS results. The near-side peak is highest in the 1.0$<\it{p}_{\rm{T}}<$2.0 interval and gradually decreases with increasing $\it{p}_{\rm{T}}$.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig2_PlotDeltaPhi.pdf}
	\caption{One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ with various transverse momentum intervals in each multiplicity class. The distributions in upper panels are measured with 0-100\% multiplicity class. The distribution in lower panels are measured with 0-0.1\% multiplicity class. Interval of transverse momentum of trigger particle and associated particle is 1.0$<\it{p}_{\rm{T}}<$2.0 GeV/\it{c}\rm{} (left), 2.0$<\it{p}_{\rm{T}}<$3.0 GeV/\it{c}\rm{} (middle) and 3.0$<\it{p}_{\rm{T}}<$4.0 GeV/\it{c}\rm{} (right), respectively. }
	\label{fig:PlotDeltaPhi}
\end{figure}

The spectra of the ridge yield are shown in Fig.~\ref{fig:PlotYSpect} in very high multiplicity class and compared with CMS results. The estimator of particle multiplicity of ALICE is done with forward subsystem(V0), whereas that of CMS is done by mid-rapidity particles meeting with the condition of $|\eta|<$2.4 and $\it{p}_{\rm{T}}>$0.4 GeV/\it{c}\rm{}. Dedicated comparison is conducted and the difference of particle multiplicity is estimated to be about 20\%. Taking into account the difference in acceptance of charged tracks and comparable definition of multiplicity,, the measurements are can be considered comparable with each other. The spectrum is also compared with model predictions. The PYTHIA8 default and PYTHIA8 + String Shoving are used for the prediction. The PYTHIA8 default is shown as black line and predicts no ridge in pp collisions, which is expected and indicates ignorable jet contamination. The PYTHIA8 + String Shoving is shown as red line and predicts existence of ridge yield in pp collisions, which decreases more rapidly than measurement as transverse momentum increases.

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.7\textwidth]{./figures/Fig1_RidgeYield.pdf} }
	\caption{ The spectra of ridge yield as function of transverse momentum. The filled circles denote the measurement with ALICE and compared with CMS measurement \cite{ridge_pp_1}, which is represented as open box. The two lines show model predictions from PYTHIA8 default(black line) and PYTHIA8 + String Shoving(red line). }
	\label{fig:PlotYSpect}
\end{figure}

@@

To further observe the behavior of the ridge yield in events including hard processes, the correlations function is measured with the minimum $\it{p}_{\rm{T, Lead}}$(left) or $\it{p}_{\rm{T, Jet}}$(right) selection for 1.0$<\it{p}_{\rm{T, assoc}},\,\it{p}_{\rm{T, trig}}<$2.0 GeV/\it{c}\rm{}, where the measured ridge yield is maximum in Fig.~\ref{fig:PlotCorrHMTSel}. The ridge structures are still seen with the event scale selection. With the minimum $\it{p}_{\rm{T, Jet}}$ selection, the correlations function has a bump on the $\Delta\eta = 0$ and $\Delta\varphi = \pi$, which occurs due to the limited jet acceptance.

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.4 \textwidth]{./figures/corrlh.pdf} }
	\subfigure{ \includegraphics[width=0.4 \textwidth]{./figures/corrjet.pdf} }
	\caption{ Two-dimensional correlations function as function of $\Delta\eta$ and $\Delta\varphi$ in top 0-0.1\% multiplicity class with the minimum $\it{p}_{\rm{T, Lead}}$(left) or $\it{p}_{\rm{T, Jet}}$(right) selection. The interval of transverse momentum of trigger particle and associated particle is 1.0$<\it{p}_{\rm{T}}<$2.0 GeV/\it{c}\rm{}. The minimum requirement for $\it{p}_{\rm{T, Lead}}$ or $\it{p}_{\rm{T, Jet}}$ is 9 GeV/\it{c}\rm{} or 10 GeV/\it{c}\rm{}, respectively. }
	\label{fig:PlotCorrHMTSel}
\end{figure}

The one-dimensional $\Delta\varphi$ distribution with the minimum $\it{p}_{\rm{T, Lead}}$(left) or $\it{p}_{\rm{T, Jet}}$ is shown in Fig.~\ref{fig:PlotDeltaPhiESE}. The near-side yield increases as the event scale increases by requiring higher $\it{p}_{\rm{T, Lead}}$(left) or $\it{p}_{\rm{T, Jet}}$. The model comparisons are seen as blue line for PYTHIA8 + String Shoving and apricot line for PYTHIA8 default, respectively. The PYTHIA8 + String Shoving describes ridge yield with the selection. The PYTHIA8 default predict the ridge yield to be zero with the selection, which also tell there is little ignorable jet contamination. Both models well describe the away-side yield with the $\it{p}_{\rm{T, Jet}}$ selection and overestimates away-side yield with the $\it{p}_{\rm{T, Lead}}$ selection.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig5_PlotDeltaPhiESE.pdf}
	\caption{ One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ with the minimum $\it{p}_{\rm{T, Lead}}$(lower) or $\it{p}_{\rm{T, Jet}}$(upper) selection for 1.0$<\it{p}_{\rm{T, assoc}},\,\it{p}_{\rm{T, trig}}<$2.0 GeV/\it{c}\rm{} in 0-0.1\% multiplicity class. The filled circles show measurement with ALICE. Model predictions are compared with measurement as blue line for PYTHIA8 + String Shoving  and apricot line for PYTHIA8 default.}
	\label{fig:PlotDeltaPhiESE}
\end{figure}

>>>>

Figure~\ref{fig:PlotCorrMBHMT} shows two-particle correlation functions as functions of the relative pseudorapidity ($\Delta \eta$) and azimuthal angle ($\Delta \varphi$) differences between the trigger and associated particles. The functions are measured with Eq.~\ref{eq:corrfunction} for $1 < \pttrig~\mathrm{and}~\ptassoc < 2$ GeV/$c$ in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV}. The left figure shows the function for the 0--0.1\% multiplicity class. The right one shows the function for the minimum bias events (the 0--100\% multiplicity class). The ridge behaviour is clearly observed in the high multiplicity class compared to the function for the minimum-bias events.

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corr1.pdf} }
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrmb.pdf} }
	\caption{ Dihadron correlation functions as functions of $\Delta\eta$ and $\Delta\varphi$ in the high multiplicity (0--0.1\%, left) and the minimum-bias events (0--100\%, right). The intervals of $\pttrig$ and $\ptassoc$ are equally $1 < \it{p}_{\rm{T}} < 2$ GeV/$c$. }
	\label{fig:PlotCorrMBHMT}
\end{figure}

 Projected $\Delta\varphi$ distributions of the two-particle correlation functions in $1.6<|\Delta\eta|<1.8$ are shown in Fig.~\ref{fig:PlotDeltaPhi} for various $\it{p}_{\rm{T}}$ intervals in the high multiplicity class (0--0.1\%). The near-side peak is highest in the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gradually decreases with increasing $\it{p}_{\rm{T}}$.
 %The associated yield per trigger particle is compared with the CMS results.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig2_PlotDeltaPhi.pdf}
	\caption{One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ with various transverse momentum intervals in each multiplicity class. The distributions in upper panels are measured with 0--100\% multiplicity class. The distribution in lower panels are measured with 0-0.1\% multiplicity class. Interval of transverse momentum of trigger particle and associated particle is $1<\pt<2$ (left), $2<\pt<3$ (middle) and $3<\pt<4$ GeV/$c$ (right), respectively. }
	\label{fig:PlotDeltaPhi}
\end{figure}

The spectra of the ridge yield are shown in Fig.~\ref{fig:PlotYSpect} in the high multiplicity class and compared with the CMS result~\cite{}. The estimator of particle multiplicity of ALICE is done with forward subsystem(V0), whereas that of CMS is done by mid-rapidity particles meeting with the condition of $|\eta|<$2.4 and $\pt>0.4$ GeV/$c$. Dedicated comparison is conducted and the difference of particle multiplicity is estimated to be about 20\%. Taking into account the difference in acceptance of charged tracks and comparable definition of multiplicity, the measurements are can be considered comparable with each other. The spectrum is also compared with model predictions. $\pythiam$ and $\pythiashoving$ are used for the prediction. $\pythiam$ is shown with a black line and predicts no ridge effect in pp collisions, which is expected because the model does not implement any constraint to describe the ridge effect. $\pythoashoving$ shown with a red line describes the effect better than $\pythiam$, however, its yield decreases more rapidly than the data as $\pttrigassoc$ increases.

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.99\textwidth]{./figures/Fig1_RidgeYield.pdf} }
	\caption{ The spectra of ridge yield as function of transverse momentum. The filled circles denote the measurement with ALICE and compared with CMS measurement \cite{ridge_pp_1}, which is represented as open box. The two lines show model predictions from $\pythiam$ (black line) and $\pythiashoving$ (red line). }
	\label{fig:PlotYSpect}
\end{figure}

@@

To further observe the behavior of the ridge yield with respect to the hard process, the correlation function is measured with the minimum $\ptlead$ (left) or $\ptjet$ (right) thresholds for $1<\pttrigassoc<2$ GeV/$c$, where the measured ridge yield is maximum in Fig.~\ref{fig:PlotCorrHMTSel}. The ridge structures are still seen with the event scale selection. With the minimum $\ptjet$ selection, the correlations function has a bump on the $\Delta\eta = 0$ and $\Delta\varphi = \pi$, which occurs due to the limited jet acceptance.

\begin{figure}[h!]
	\centering
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrlh.pdf} }
	\subfigure{ \includegraphics[width=0.47 \textwidth]{./figures/corrjet.pdf} }
	\caption{ Two-dimensional correlations function as function of $\Delta\eta$ and $\Delta\varphi$ in top 0-0.1\% multiplicity class with the minimum $\ptlead$ (left) or $\ptjet$ (right) selection. The interval of $\pttrig$ and $\ptassoc$ is $1<\pt<2$ GeV/$c$. The minimum requirement for $\ptlead$ or $\ptjet$ is 9 and 10 GeV/$c$, respectively. }
	\label{fig:PlotCorrHMTSel}
\end{figure}

The one-dimensional $\Delta\varphi$ distribution with the minimum $\ptlead$ (left) or $\ptjet$ is shown in Fig.~\ref{fig:PlotDeltaPhiESE}. The near-side yield increases as the event scale increases by requiring higher $\ptlead$ (left) or $\ptjet$. The model comparisons are are shown for $\pythiashoving$ with a blue and $\pythiam$ with a red line. The $\pythiashoving$ describes ridge yield with the selection. $\pythiam$ predicts non ridge yield with the selection, which also tell there is little ignore-able jet contamination. Both models well describe the away-side yield with the $\ptjet$ selection and overestimates away-side yield with the $\ptlead$ selection.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig5_PlotDeltaPhiESE.pdf}
	\caption{ One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ with the minimum $\ptlead$(lower) or $\ptjet$(upper) selection for $1<\pttrigassoc<2$ GeV/\it{c}\rm{} in the 0--0.1\% multiplicity class. The filled circles show measurement with ALICE. Model predictions are compared with measurement with a blue line for $\pythiashoving$  and a orange line for $\pythiam$.}
	\label{fig:PlotDeltaPhiESE}
\end{figure}

### 2020-06-12 sec_summary.tex
JL and BK lastet

ALICE has measured the $p_{\mathrm T}$ and the $\it{p}_{\rm{T,Lead}}$ or $\it{p}_{\rm{T,Jet}}$ selection dependence of the long-range azimuthal correlations in very high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV at the Large Hadron Collider. The spectrum of the ridge yield has been measured in the relative pseudorapidity range $1.6 < |\eta| < 1.8$ for the transverse momentum range $1.0 < p_{\mathrm T} < 5.0$~GeV/$c$ and 0--0.1\% multiplicity event class. The result has been compared with CMS measurement and shown that the results are compatible. Further understandings of the correlations in small system have been tried by studying the ridge yield in events, where jets with harder fragmentations is to be present by requiring the $\it{p}_{\rm{T,Lead}}$ or $\it{p}_{\rm{T,Jet}}$. The ridge yield spectrum with hard fragmentations($p_{\mathrm T, jet} > 10$~GeV/$c$.) shows enhancement as the $\it{p}_{\rm{T,Lead}}$ or $\it{p}_{\rm{T,Jet}}$ increases. The results have been compared with PYTHIA8 string shoving calculation, showing comparable enhancement as data.

>>>>

%Two-particle angular correlations in a large pseudorapidity difference of $1.6<|\Delta\eta|<1.8$ are measured in very high-multiplicity pp collisions at $\sqrt{\it{s}}=13$ TeV with ALICE. The measured associated yield is found to be consistent with the previous results from CMS experiment. The ridge yield, for the first time, is observed with respect to the hard process. Furthermore, it is found to be independent of the hardness of events. This observation is important for the study of the origins of the ridge in small collision systems.

Ridge yields via long-range azimuthal correlations are measured with respect to the hard process in pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV.
The hard process is selected respectively by the transverse momenta of the leading track $p_\mathrm{T,\,Lead}$ and leading jet $p_\mathrm{T,\,Jet}$ for each event. $\p_\mathrm{T,\,Lead}$ is measured in a pseudorapidity region of $|\eta|<0.9$ for the full azimuthal angle. Jets are reconstructed with charged particles using the anti-$k_\mathrm{T}$ algorithm for a resolution parameter $R=0.4$ in a pseudorapidity region of $|\eta_\mathrm{jet}|<0.4$ for the full azimuthal angle. 
The ridge yield is measured in the relative pseudorapidity range $1.6 < |\Delta\eta| < 1.8$ for the transverse momentum range $1 < p_\mathrm{T} < 5$ GeV/$c$. The high multiplicity pp collisions are selected in terms of the energy deposit in the V0 detector of ALICE sensitive to charged-particle multiplicity in the forward region.

The ridge behaviour is clearly appeared in high-multiplicity pp collisions that are relevant to the top 0.1\% events for the energy deposit in the V0. The result is compared to the CMS result and it is found to be compatible. The PYTHIA 8 String Shoving estimates the ridge yield quantitatively while PYTHIA 8 does not describe the ridge effect at all. Further study of the ridge behaviors in small system is done by measuring a ridge yield in events, where jets accompanying with the hard scattering process present by requiring thresholds for $p_\mathrm{T,\,Lead}$ and $p_\mathrm{T,\,Jet}$. The ridge yield for events with $p_\mathrm{T,\,jet} > 10$ GeV/$c$ is increased as $p_\mathrm{T,\,Lead}$ and $p_\mathrm{T,\,Jet}$ increase. The results are compared with PYTHIA 8 String Shoving model. The model describes the enhancement the ridge yield, however, underestimating the absolute ridge yield than data. This observation is important for the study of the origins of the ridge in small systems.
